### 2022-03-04 paper/localpsf.tex
shortening and simplifying radial basis function writing

Having found the $k$ nearest valid sample points $x_i$ to $x$, we approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
\begin{equation}
\widetilde{\Phi}(y,x) := \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right),
\end{equation}
where $b_i$ are radial basis functions, and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
\begin{equation}
Bw = f,
\end{equation}
where the coefficient matrix $B \in \mathbb{R}^{k \times k}$ has entries 
\begin{equation*}
B_{ij} := b\left(\|x_i - x_j\|\right), \qquad i,j=1,\dots,k,
\end{equation*}
and the right hand side vector $f \in \mathbb{R}^k$ has entries 
\begin{equation*}
f_i = f_{y,x}(x_i), \qquad i=1,\dots,k.
\end{equation*}
We use polyharmonic spline radial basis functions \cite{polyspline}, which are defined as follows:
\begin{equation}
b_i(r) := \begin{cases}
r^d, & d=1,3,5,\dots \\
r^d \log r & d=2,4,6,\dots
\end{cases}
\end{equation}
for $r>0$, and $b_i(0):=0$.
%\begin{equation}
%	f_i = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
%\end{equation}
%The result of this interpolation is to approximate $\Phi(y,x)$ with a linear combination of the shifted impulse response values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$. 
%By including the factor $\frac{V(x)}{V(x_i)}$, we are implicitly interpolating the ``volume preconditioned'' function $\frac{1}{V(x)}\Phi(y,x)$, then rescaling the result by $V(x)$ to get an approximation of $\Phi(y,x)$. Volume preconditioning improves the approximation considerably when $V(x)$ varies widely in magnitude, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.

@@

To generate the right hand side vector $f$, we evaluate $f_{y,x}(x_i)$ at the points $x_i$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
\begin{equation}
\label{eq:ellipsoid_check93}
(y - \mu(x))^T \Sigma^{-1}(x) (y - \mu(x))  \le \tau^2.
\end{equation}
If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i) \notin E_{x_i}$ (i.e., $y - \mu(x) + \mu(x_i)$ is outside the estimated support of $\impulseresponse_{x_i}$), so we set $f_{y,x}(x_i)=0$ for the purpose of interpolation. If \eqref{eq:ellipsoid_check93} does hold, then $y - \mu(x) + \mu(x_i) \in E_{x_i}$. In this case, we look up the batch index $b$ for the batch of sample points that contains $x_i$, and evaluate
\begin{equation}
f_{y,x}(x_i) = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right) = V(x)\eta_b\left(y - \mu(x) + \mu(x_i)\right).
\end{equation}
%Recall from Section \ref{sec:get_impulse_response} that $\eta_b$ is the impulse response batch that we have computed and stored.
Evaluating the impulse response batch $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires some care, because the point $y - \mu(x) + \mu(x_i)$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
To evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, we determine which mesh cell contains $y - \mu(x) + \mu(x_i)$, then we evaluate the relevant finite element basis functions on that mesh cell. Fortunately, the mesh cell containing $y - \mu(x) + \mu(x_i)$ was already determined as a side effect of querying the AABB tree of mesh cells when we checked whether $y - \mu(x) + \mu(x_i) \in \Omega$.

>>>>

Having found the $k$ nearest valid sample points $x_i$ to $x$, we approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
\begin{equation}
\widetilde{\Phi}(y,x) := \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right),
\end{equation}
where $b_i$ are radial basis functions, and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
\begin{equation}
Bw = f,
\end{equation}
where the coefficient matrix $B \in \mathbb{R}^{k \times k}$ and right hand side vector $f \in \mathbb{R}^k$ have entries 
\begin{align*}
B_{ij} :=& b\left(\|x_i - x_j\|\right),\\
f_i :=& f_{y,x}(x_i),
\end{align*}
for $i,j=1,\dots,k$.
% has entries 
%\begin{equation*}
%f_i = f_{y,x}(x_i), \qquad i=1,\dots,k.
%\end{equation*}
We use polyharmonic spline radial basis functions \cite{polyspline}, which are defined as follows:
\begin{equation}
b_i(r) := \begin{cases}
r^d, & d=1,3,5,\dots \\
r^d \log r & d=2,4,6,\dots
\end{cases}
\end{equation}
for $r>0$, and $b_i(0):=0$.
%\begin{equation}
%	f_i = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
%\end{equation}
%The result of this interpolation is to approximate $\Phi(y,x)$ with a linear combination of the shifted impulse response values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$. 
%By including the factor $\frac{V(x)}{V(x_i)}$, we are implicitly interpolating the ``volume preconditioned'' function $\frac{1}{V(x)}\Phi(y,x)$, then rescaling the result by $V(x)$ to get an approximation of $\Phi(y,x)$. Volume preconditioning improves the approximation considerably when $V(x)$ varies widely in magnitude, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.

@@

We evaluate $f_{y,x}(x_i)$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
\begin{equation}
\label{eq:ellipsoid_check93}
(y - \mu(x))^T \Sigma^{-1}(x) (y - \mu(x))  \le \tau^2.
\end{equation}
If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i)$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_{y,x}(x_i)=0$ for the purpose of interpolation. If \eqref{eq:ellipsoid_check93} does hold, then we look up the batch index $b$ for such that $x_i \in S_b$, and evaluate
\begin{equation}
f_{y,x}(x_i) = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right) = V(x)\eta_b\left(y - \mu(x) + \mu(x_i)\right).
\end{equation}
%Recall from Section \ref{sec:get_impulse_response} that $\eta_b$ is the impulse response batch that we have computed and stored.
%Evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires care, because 
When the problem is discretized with finite elements, $y - \mu(x) + \mu(x_i)$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
To evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, we therefore determine which mesh cell contains $y - \mu(x) + \mu(x_i)$, then we evaluate the relevant finite element basis functions on that mesh cell. Fortunately, the mesh cell containing $y - \mu(x) + \mu(x_i)$ was already determined as a side effect of querying the AABB tree of mesh cells when we checked whether $y - \mu(x) + \mu(x_i) \in \Omega$.
